Unimodular covers of 3-dimensional parallelepipeds and Cayley sums

### 2021-02-25 unimodular-covers.tex
updated Paco's grant info
comment on corollary

\thanks{The authors were supported by the Einstein Foundation Berlin under grant EVF-2015-230. 
Work of F. Santos is also supported by project MTM2017-83750-P of the Spanish Ministry of Science (AEI/FEDER, UE)}

@@

\begin{abstract}
We show that the following classes of lattice polytopes have unimodular covers, in dimension three: the class of parallelepipeds, the class of centrally symmetric polytopes, and the class of Cayley sums $\cayley(P,Q)$ where the normal fan of $Q$ refines that of $P$. This improves results of Beck et al.~(2018) and Haase et al.~(2008) where the last two classes were shown to be IDP.
\end{abstract}

@@

@@

\begin{question}
Does every smooth 3-polytope have the circumscribed parallelepiped property? 
\end{question}

>>>>

\thanks{The authors were supported by the Einstein Foundation Berlin under grant EVF-2015-230. 
Work of F. Santos is also supported by grants  MTM2017-83750-P/AEI/10.13039/501100011033 and PID2019-106188GB-I00/AEI/10.13039/501100011033 of the Spanish State Research Agency}

@@

\begin{abstract}
We show that the following classes of lattice polytopes have unimodular covers, in dimension three:  parallelepipeds,  smooth centrally symmetric polytopes, and  Cayley sums $\cayley(P,Q)$ where the normal fan of $Q$ refines that of $P$. This improves results of Beck et al.~(2018) and Haase et al.~(2008) where the last two classes were shown to be IDP.
\end{abstract}
\paco{added ``smooth'' to c.s. polytopes in abstract}

@@

\Paco{
I am a bit puzzled. After the next corollary we say that \cite[Cor.~4.2]{SantosZiegler} proves a case of this, but in fact it proves the whole of it, in a stronger form that does not need width one, and that gives a "dissection" and not a mere cover (dissection= cover by simplices with disjoint interiors).

I propose to change it to the following, related to a recent preprint of Gubeladze:

\noindent{\bf Corollary}
%\label{coro:gorenstein}
Let $P$ be a lattice $3$-polytope with the following property: for every facet $F$ of $P$ and every edge $E$ of $P$ with (exactly) an end-point in $F$, there is a lattice point $v\in E$ at distance $1$ from $P$. Then, there is a unimodular cover of its ``lattice boundary''. 
%\end{corollary}

Here ee call ``lattice boundary'' of a polytope the union of all the prismatoids obtained translating each facet hyperplane one lattice unit towards the interior. Put differently, if $\{a_i x \le b_i, i=1,\dots,m\}$ integer is an integral facet description of $P$, the lattice boundary is $P$ minus the interior of $P'=\{a_i x \le b_i-1, i=1,\dots,m\}$. Remark: $P'$ is related to (but not exactly the same as) the ``Fine interior'' of $P$. See eg, some papers by Batyrev on this.

This corollary is related (but neither implies nor is implied by) Theorem (a) in Gubeladze's recent arXiv:2012.11846
}

@@

\begin{question}
Does every smooth 3-polytope have the circumscribed parallelepiped property? 
\end{question}
\Paco{
The following is a counter-example: Let $P$ be the Cayley embedding of a long horizontal rectangle and a long vertical rectangle. That is,
$P=conv([0,a]\times [0,1] \times\{0\} \cup [0,1]\times [0,b] \times\{1\})$, for big $a$ and $b$. This is smooth and contains the big empty tetrahedron $T$ with vertices (0,0,0), (a,1,0), (0,0,1), (1,b,1) which occupies most of its volume. In particular, none of the corner tetrahedra of $T$ is contained in $P$.

A weaker version of the question is:

Question 2.5Õ. If $T$ is an empty tetrahedron contained in a smooth 3-polytope $P$, can one guarantee that there is a lattice point of $P$ in the circumscribed parallelepiped of $T$ (apart of the vertices of $T$)?

A positive answer would still imply that $P$ can be unimodularly covered.
}
